---
title: "AppS1_17Nov2023"
author: "Madison McCaig"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
- run the main manuscript rmd before this one. 

# Fig S2 and S3: Rarefaction Curves

- can be found in the QA/QC files

# Fig S4: Stream Habitat Correlations (Full Season)

```{r Correlations with Cumulative Defoliation, fig.height=12}
#join defo and waterchem means
cumulative_defo_waterchem <- waterchem_means_wide %>%
        inner_join(HAIFLO, by = c("site", "year"))

cumulative_defo_waterchem <- cumulative_defo_waterchem %>%
  dplyr::select(-3)

#add the DOM data and flow data and temp data

cumulative_defo_waterchem <- cumulative_defo_waterchem %>%
        inner_join(DOM_data_summary, by = c("site", "year"))

cumulative_defo_waterchem <- cumulative_defo_waterchem %>%
        inner_join(flow_fullseason_means, by = c("site", "year"))

cumulative_defo_waterchem <- cumulative_defo_waterchem %>%
        inner_join(temp_levellogger_summary, by = c("site", "year"))
  
cumulative_defo_waterchem$site <- as.factor(cumulative_defo_waterchem$site)
cumulative_defo_waterchem$year <- as.factor(cumulative_defo_waterchem$year)

# then restrict to only the ones we predicted to change
cdefo_stream <- cumulative_defo_waterchem %>%
  dplyr::select(-3, -4, -5, -6, -7, -8, -9, -12, -13, -14, -15, -17, -18, -19, -21, -23, -24, -31)
#rename everything
cdefo_stream <- cdefo_stream %>%
  dplyr::rename(SUVA254 = mean_SUVA, flow.m3s = flow.m3s_mean, temp.C = temp_mean, NO2NO3 = NO2.NO3.mgL)

#now need to run correlations of each waterchem parameter and each year separate

cdefo_stream_2019 <- cdefo_stream  %>%
  filter(year == "2019") %>%
   pivot_longer(cols= -c("site", "cdefo","year"), names_to="parameter", values_to="param_val")

cdefo_stream_2020 <- cdefo_stream  %>%
  filter(year == "2020") %>%
   pivot_longer(cols= -c("site", "cdefo","year"), names_to="parameter", values_to="param_val")
   
cdefo_stream_2021 <- cdefo_stream  %>%
  filter(year == "2021") %>%
   pivot_longer(cols= -c("site", "cdefo","year"), names_to="parameter", values_to="param_val")

library(ggpubr)
p1 <- cdefo_stream_2019 %>%
ggscatter(x = 'cdefo', y = 'param_val',
         add = 'reg.line',
         conf.int = TRUE,
         cor.coef = TRUE,
         cor.method = 'pearson',
         xlab = "Cumulative Defoliation Index value (0-15)",
         ylab = "Value of stream habitat parameter") +
facet_wrap(~parameter, scales = "free_y") +
theme_bw()

p2 <- cdefo_stream_2020 %>%
ggscatter(x = 'cdefo', y = 'param_val',
         add = 'reg.line',
         conf.int = TRUE,
         cor.coef = TRUE,
         cor.method = 'pearson',
         xlab = "Cumulative Defoliation Index value (0-15)",
         ylab = "Value of stream habitat parameter") +
facet_wrap(~parameter, scales = "free_y") +
theme_bw()

p3 <- cdefo_stream_2021 %>%
ggscatter(x = 'cdefo', y = 'param_val',
         add = 'reg.line',
         conf.int = TRUE,
         cor.coef = TRUE,
         cor.method = 'pearson', 
         xlab = "Cumulative Defoliation Index value (0-15)",
         ylab = "Value of stream habitat parameter") +
facet_wrap(~parameter, scales = "free_y") +
theme_bw()

#try everything together and facet by year
cdefo_stream_test <- cdefo_stream  %>%
   pivot_longer(cols= -c("site", "cdefo","year"), names_to="parameter", values_to="param_val")

cdefo_stream_test %>%
ggscatter(x = 'cdefo', y = 'param_val',
         add = 'reg.line',
         conf.int = TRUE,
         cor.coef = TRUE,
         cor.coef.size = 5,
         cor.method = 'pearson', 
         xlab = "Cumulative Defoliation Index value (0-15)",
         ylab = "Value of stream habitat parameter") +
facet_grid(parameter~year, scales = "free_y") +
theme_bw(base_size = 18)

```

# Fig S5: Cumulative Precipiation 

- data from the New Richmond Est Environment Canada Weather Station https://climate.weather.gc.ca/historical_data/search_historic_data_e.html 
- graph showing cumulative precipitation from Nov 1 of the previous year to Oct 31 of that year
- i.e. 2019 is Nov 1 2018 to Oct 31 2019
- this is to account of snow accumulation/spring melt in the streams

```{r}
library(readr)
weather_2018 <- read_csv("~/SBW Data/EnviroCan Weather/en_climate_daily_QC_7055440_2018_P1D.csv")

weather_2018 <- weather_2018 %>%
  dplyr::select(5:8, 10,12,14,16,18, 24,26)

colnames(weather_2018)[10] ="Precip_mm"

weather_2019 <- read_csv("~/SBW Data/EnviroCan Weather/en_climate_daily_QC_7055440_2019_P1D.csv")
weather_2019 <- weather_2019 %>%
  dplyr::select(5:8, 10,12,14,16,18, 24,26)
colnames(weather_2019)[10] ="Precip_mm"

weather_2020 <- read_csv("~/SBW Data/EnviroCan Weather/en_climate_daily_QC_7055440_2020_P1D.csv")
weather_2020 <- weather_2020 %>%
  dplyr::select(5:8, 10,12,14,16,18, 24,26)
colnames(weather_2020)[10] ="Precip_mm"

weather_2021 <- read_csv("~/SBW Data/EnviroCan Weather/en_climate_daily_QC_7055440_2021_P1D.csv")
weather_2021 <- weather_2021 %>%
  dplyr::select(5:8, 10,12,14,16,18, 24,26)
colnames(weather_2021)[10] ="Precip_mm"

precip2019 <- rbind(weather_2018[305:365,], weather_2019[1:304,])
precip2019$Precip_mm[is.na(precip2019$Precip_mm)] <- 0

precip2020 <- rbind(weather_2019[305:365,], weather_2020[1:305,])
precip2020$Precip_mm[is.na(precip2020$Precip_mm)] <- 0

precip2021 <- rbind(weather_2020[306:366,], weather_2021[1:304,])
precip2021$Precip_mm[is.na(precip2021$Precip_mm)] <- 0

precip2019$precip_cumulative <- cumsum(precip2019$Precip_mm)
precip2020$precip_cumulative <- cumsum(precip2020$Precip_mm)
precip2021$precip_cumulative <- cumsum(precip2021$Precip_mm)

precip2019$Year <- "2019"
precip2020$Year <- "2020"
precip2021$Year <- "2021"

library(tibble)
precip2019 <- tibble::rownames_to_column(precip2019, "DayNumber")
precip2019$DayNumber <- as.numeric(precip2019$DayNumber)
precip2020 <- tibble::rownames_to_column(precip2020, "DayNumber")
precip2020$DayNumber <- as.numeric(precip2020$DayNumber)
precip2021 <- tibble::rownames_to_column(precip2021, "DayNumber")
precip2021$DayNumber <- as.numeric(precip2021$DayNumber)

precip_final <- rbind(precip2019, precip2020, precip2021)

precip_final$Year <- as.factor(precip_final$Year)

library(scales)
p <- ggplot(precip_final, aes(x = as.Date(DayNumber, "1970-11-01"), y = precip_cumulative, color = Year)) +
  geom_line(size=2) +
  scale_x_date(date_breaks="months", date_labels="%b") +
  labs(x="Month",colour="")  +
  theme_bw(base_size =16) + ylab("Cumulative Precipitation (mm)") 

p + theme(legend.position = "bottom")
```

# Fig S6 Stream Flow Models

```{r flow data}

defo_flow <- flow_fullseason_means %>%
        inner_join(HAIFLO, by = c("site", "year"))

#add a column for Region (lower, central, upper)
defo_flow <- defo_flow %>%
  mutate(Region = case_when(
    startsWith(site, "L") ~ "Lower",
    startsWith(site, "C") ~ "Central",
    startsWith(site, "U") ~ "Upper"
    ))

p1 <- defo_flow  %>%
        ggscatter(x = "cdefo", y = "flow.m3s_mean", color = "year", shape = "Region", 
                  size = 5, add = "reg.line", conf.int = TRUE,
                  show.legend.text = FALSE) + stat_cor(p.accuracy = 0.01, r.accuracy = 0.01, aes(size =6, label = paste(..r.label.., ..p.label.., sep = "~`,`~")), show.legend = FALSE, size = 6) + scale_color_brewer(palette = "Dark2") + scale_fill_brewer(palette = "Dark2", guide = "none") + facet_wrap(~year) +xlab("Cumulative Defoliation") + ylab(bquote('Stream Flow ('*m^3/s*')')) + theme_bw(base_size = 22) + theme(legend.position = "bottom", panel.grid.major = element_blank(),
panel.grid.minor = element_blank()) + theme(legend.box.background=element_rect(color = "black", size = 1))


mydata <- defo_flow %>%
  dplyr::rename(flow = flow.m3s_mean)

#SUVA_HAiFLO <- readRDS("C:/Users/madis/OneDrive - McMaster University/SBWD_data_analysis/EditedData/DOM/SUVA_HAiFLO.RDS")

#install.packages("readxl")
library(readxl)
watersheds <- read_excel(path="~/SBW Data/watersheds.xlsx" ,
                  sheet="watersheds",
                  col_names= TRUE)

AnnualHydro <- AnnualHydro %>%
        rename(site = Site) %>%
        rename(year = Year)

AnnualHydro$year <- as.factor(AnnualHydro$year)

mydata <- mydata %>%
        inner_join(AnnualHydro, by = c("site", "year"))

watersheds <- watersheds %>%
        rename(site = Site)

mydata <- mydata %>%
        inner_join(watersheds, by = c("site"))

watersheds_mortality_statistics <- readRDS("~/SBW Data/SBWD_master-main_2023May15/SBWD_master-main/data/watersheds_mortality_statistics.RDS")

mortality <- watersheds_mortality_statistics %>%
  filter(hydroweighting == "HAiFLO") %>%
  dplyr::select(1,3,6) %>%
  dplyr::rename(Mortality = value)

mydata <- mydata %>%
  left_join(mortality, by = c("site", "year"))
#change NAs to 0
mydata <- mydata %>%
  replace_na(list(Mortality = 0))

mydata19 <- mydata %>%
        filter(year=="2019")

mydata20 <- mydata %>%
        filter(year=="2020")

mydata21 <- mydata %>%
        filter(year=="2021")
```

```{r flow HP models, results = 'hide', fig.width=12, fig.height=14}
#elevation, slope, defo, spfir, catchment area
HPmod_land19 <- rdacca.hp(mydata19$flow, mydata19[,c("Mean_Elevation", "cdefo", "Catchment_Area_km2", "SpFir_Percent", "Mean_slope_degrees")], var.part = TRUE)
#plot(HPmod_land19, plot.perc = TRUE)
HPmod_land19

HPmod_land20 <- rdacca.hp(mydata20$flow, mydata20[,c("Mean_Elevation", "cdefo", "Catchment_Area_km2", "SpFir_Percent", "Mean_slope_degrees")], var.part = TRUE)
#plot(HPmod_land20, plot.perc = TRUE)
HPmod_land20

HPmod_land21<- rdacca.hp(mydata21$flow, mydata21[,c("Mean_Elevation", "cdefo", "Catchment_Area_km2", "SpFir_Percent", "Mean_slope_degrees")], var.part = TRUE)
#plot(HPmod_land21, plot.perc = TRUE)
HPmod_land21

#plot total and then put unique and shared into a table

HP_vals_c21 <- (as.data.frame(HPmod_land21[[4]]))
HP_vals_c21$parameter <- row.names(HP_vals_c21)  

HP_vals_c21 <- HP_vals_c21 %>%
dplyr::select(4,5)

HP_vals_c21_long <- HP_vals_c21 
       # %>% pivot_longer(
             #   cols = 1:2,
               # names_to = c("condition"),
               # values_to = "value")

HP_vals_c21_long$year <- "2021"
 

#2019
HP_vals_c19 <- (as.data.frame(HPmod_land19[[4]]))
HP_vals_c19$parameter <- row.names(HP_vals_c19)  

HP_vals_c19 <- HP_vals_c19 %>%
dplyr::select(4,5)

HP_vals_c19_long <- HP_vals_c19 
#%>% 
        #pivot_longer(
              #  cols = 1:2,
              #  names_to = c("condition"),
              #  values_to = "value")

HP_vals_c19_long$year <- "2019"
#2020
HP_vals_c20 <- (as.data.frame(HPmod_land20[[4]]))
HP_vals_c20$parameter <- row.names(HP_vals_c20)  

HP_vals_c20 <- HP_vals_c20 %>%
dplyr::select(4,5)

HP_vals_c20_long <- HP_vals_c20 
#%>% 
      #  pivot_longer(
            #    cols = 1:2,
             #   names_to = c("condition"),
             #   values_to = "value")
HP_vals_c20_long$year <- "2020"
                                    
#join years together
HP_vals_long <- rbind(HP_vals_c19_long, HP_vals_c20_long, HP_vals_c21_long)

#HP_vals_long <- HP_vals_long %>%
  #dplyr::rename(Condition = condition)

HP_vals_long <- HP_vals_long %>%
dplyr::rename(Indiv.perc = "I.perc(%)")
#HP_vals_long$Condition <- as.factor(HP_vals_long$Condition)

#need to change order of the factors
HP_vals_long$parameter <- HP_vals_long$parameter %>%
factor(levels = c("Mean_Elevation", "cdefo", "SpFir_Percent", "Mean_slope_degrees", "Catchment_Area_km2"))

p2 <- ggplot(HP_vals_long, aes(y=Indiv.perc, x=parameter, fill = parameter)) + 
    geom_bar(stat="identity") + facet_wrap(~year) +   
 scale_x_discrete(labels=c("Elevation", "Cumulative Defoliation","% Spruce-Fir", "Slope", "Catchment Area")) + 
  theme_bw(base_size = 22) +
 xlab("Landscape Metric") + ylab(bquote('% Individual Contribution to '*R^2*'')) +
#scale_fill_discrete(labels=c('Total Variance', 'Unique Variance')) +
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.border = element_rect(color="black", fill=NA), axis.text.x = element_text(angle = 45, hjust = 1, vjust=1), legend.position = "none") + scale_fill_brewer(palette="Dark2")

#1000 by 1400
library(ggpubr)
ggarrange(p1, p2,
          widths = 7,
          labels = c("a", "b"),
          font.label=list(size = 22, color = "black", face = "bold"),
          ncol = 1, nrow = 2)
```

# Fig S7 Stream Temperature (Yearly Models)

## Temperature - Incubation period

```{r temp incubation, results = 'hide', fig.width=12, fig.height=14}
dailytemp_summary <- readRDS("~/SBWD_data_analysis/EditedData/StreamHabitat/dailytemp_summary.RDS")

#does cdefo correlate to temp?
defo_temp <- dailytemp_summary %>%
        inner_join(HAIFLO, by = c("site", "year"))

#add a column for Region (lower, central, upper)
defo_temp <- defo_temp %>%
  mutate(Region = case_when(
    startsWith(site, "L") ~ "Lower",
    startsWith(site, "C") ~ "Central",
    startsWith(site, "U") ~ "Upper"
    ))

defo_temp <- defo_temp %>%
  dplyr::rename(temp_mean = mean)

#plot by year
p1 <- defo_temp  %>%
        ggscatter(x = "cdefo", y = "temp_mean", color = "year", shape = "Region", 
                  size = 5, add = "reg.line", conf.int = TRUE,
                  show.legend.text = FALSE) + stat_cor(p.accuracy = 0.01, r.accuracy = 0.01, aes(size =6, label = paste(..r.label.., ..p.label.., sep = "~`,`~")), show.legend = FALSE, size = 6) + scale_color_brewer(palette = "Dark2") + scale_fill_brewer(palette = "Dark2", guide = "none") + facet_wrap(~year) +xlab("Cumulative Defoliation") + ylab(bquote(paste("Stream Temperature", " (", "\u00B0", "C", ")", sep = ""))) + theme_bw(base_size = 22) + theme(legend.position = "bottom", panel.grid.major = element_blank(),
panel.grid.minor = element_blank()) + theme(legend.box.background=element_rect(color = "black", size = 1))

mydata <- defo_temp

AnnualHydro <- readRDS("~/SBWD_data_analysis/EditedData/DischargeSummary_Aug2022.RDS")
#SUVA_HAiFLO <- readRDS("C:/Users/madis/OneDrive - McMaster University/SBWD_data_analysis/EditedData/DOM/SUVA_HAiFLO.RDS")

#install.packages("readxl")
library(readxl)
watersheds <- read_excel(path="~/SBW Data/watersheds.xlsx" ,
                  sheet="watersheds",
                  col_names= TRUE)

AnnualHydro <- AnnualHydro %>%
        rename(site = Site) %>%
        rename(year = Year)

AnnualHydro$year <- as.factor(AnnualHydro$year)

mydata <- mydata %>%
        inner_join(AnnualHydro, by = c("site", "year"))

watersheds <- watersheds %>%
        rename(site = Site)

mydata <- mydata %>%
        inner_join(watersheds, by = c("site"))

watersheds_mortality_statistics <- readRDS("~/SBW Data/SBWD_master-main_2023May15/SBWD_master-main/data/watersheds_mortality_statistics.RDS")

mortality <- watersheds_mortality_statistics %>%
  filter(hydroweighting == "HAiFLO") %>%
  dplyr::select(1,3,6) %>%
  dplyr::rename(Mortality = value)

mydata <- mydata %>%
  left_join(mortality, by = c("site", "year"))
#change NAs to 0
mydata <- mydata %>%
  replace_na(list(Mortality = 0))

mydata19 <- mydata %>%
        filter(year=="2019")

mydata20 <- mydata %>%
        filter(year=="2020")

mydata21 <- mydata %>%
        filter(year=="2021")

#elevation, defo, spfir, aspect, wet cover
HPmod_land19 <- rdacca.hp(mydata19$temp_mean, mydata19[,c("Mean_Elevation", "cdefo", "SpFir_Percent", "Mean_aspect_degrees", "Wetland_PercentCover")], var.part = TRUE)
#plot(HPmod_land19, plot.perc = TRUE)
HPmod_land19

HPmod_land20 <- rdacca.hp(mydata20$temp_mean, mydata20[,c("Mean_Elevation", "cdefo", "SpFir_Percent", "Mean_aspect_degrees", "Wetland_PercentCover")], var.part = TRUE)
#plot(HPmod_land20, plot.perc = TRUE)
HPmod_land20

HPmod_land21<- rdacca.hp(mydata21$temp_mean, mydata21[,c("Mean_Elevation", "cdefo", "SpFir_Percent", "Mean_aspect_degrees", "Wetland_PercentCover")], var.part = TRUE)
#plot(HPmod_land21, plot.perc = TRUE)
HPmod_land21

#plot total and then put unique and shared into a table

HP_vals_c21 <- (as.data.frame(HPmod_land21[[4]]))
HP_vals_c21$parameter <- row.names(HP_vals_c21)  

HP_vals_c21 <- HP_vals_c21 %>%
dplyr::select(4,5)

HP_vals_c21_long <- HP_vals_c21 
       # %>% pivot_longer(
             #   cols = 1:2,
               # names_to = c("condition"),
               # values_to = "value")

HP_vals_c21_long$year <- "2021"
 

#2019
HP_vals_c19 <- (as.data.frame(HPmod_land19[[4]]))
HP_vals_c19$parameter <- row.names(HP_vals_c19)  

HP_vals_c19 <- HP_vals_c19 %>%
dplyr::select(4,5)

HP_vals_c19_long <- HP_vals_c19 
#%>% 
        #pivot_longer(
              #  cols = 1:2,
              #  names_to = c("condition"),
              #  values_to = "value")

HP_vals_c19_long$year <- "2019"
#2020
HP_vals_c20 <- (as.data.frame(HPmod_land20[[4]]))
HP_vals_c20$parameter <- row.names(HP_vals_c20)  

HP_vals_c20 <- HP_vals_c20 %>%
dplyr::select(4,5)

HP_vals_c20_long <- HP_vals_c20 
#%>% 
      #  pivot_longer(
            #    cols = 1:2,
             #   names_to = c("condition"),
             #   values_to = "value")
HP_vals_c20_long$year <- "2020"
                                    
#join years together
HP_vals_long <- rbind(HP_vals_c19_long, HP_vals_c20_long, HP_vals_c21_long)

#HP_vals_long <- HP_vals_long %>%
  #dplyr::rename(Condition = condition)

HP_vals_long <- HP_vals_long %>%
dplyr::rename(Indiv.perc = "I.perc(%)")
#HP_vals_long$Condition <- as.factor(HP_vals_long$Condition)

#need to change order of the factors
HP_vals_long$parameter <- HP_vals_long$parameter %>%
factor(levels = c("SpFir_Percent","cdefo", "Mean_Elevation","Wetland_PercentCover","Mean_aspect_degrees"))

p2 <- ggplot(HP_vals_long, aes(y=Indiv.perc, x=parameter, fill = parameter)) + 
    geom_bar(stat="identity") + facet_wrap(~year) +   
 scale_x_discrete(labels=c("% Spruce-Fir", "Cumulative Defoliation", "Elevation", "% Wetland", "Aspect")) + 
  theme_bw(base_size = 22) +
 xlab("Landscape Metric") + ylab(bquote('% Individual Contribution to'*R^2*'')) +
#scale_fill_discrete(labels=c('Total Variance', 'Unique Variance')) +
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"), panel.border = element_rect(color="black", fill=NA), axis.text.x = element_text(angle = 45, hjust = 1, vjust=1), legend.position = "none") + scale_fill_brewer(palette="Dark2")

library(ggpubr)
ggarrange(p1, p2,
          labels = c("A", "B"),
          font.label=list(size = 22, color = "black", face = "bold"),
          ncol = 1, nrow = 2)
```

# Fig S8 Stream Temperature (Regional Models)

```{r}
#plot by region only
p3 <- defo_temp  %>%
        ggscatter(x = "cdefo", y = "temp_mean", color = "year", 
                  size = 5, conf.int = FALSE,
                  show.legend.text = FALSE) + stat_cor(p.accuracy = 0.01, r.accuracy = 0.01, aes(size =6, label = paste(..r.label.., ..p.label.., sep = "~`,`~")), show.legend = FALSE, size = 6) + scale_color_brewer(palette = "Dark2") + scale_fill_brewer(palette = "Dark2", guide = "none") + facet_grid(~Region) +xlab("Cumulative Defoliation") + ylab(bquote(paste("Stream Temperature", " (", "\u00B0", "C", ")", sep = ""))) + theme_bw(base_size = 22) + theme(legend.position = "bottom", panel.grid.major = element_blank(),
panel.grid.minor = element_blank()) + theme(legend.box.background=element_rect(color = "black", size = 1))

p4 <- p3 + geom_smooth(method = "lm", color = "black")
```

# Fig S9 Alpha Diversity

```{r alpha diversity, fig.width=12, fig.height=12}
diversity_its <- readRDS("~/SBWD_data_analysis/EditedData/Communities/diversity/diversity_its.RDS")
lp_its_richness <- readRDS("~/SBWD_data_analysis/EditedData/Communities/diversity/lp_its_richness.RDS")
diversity_16s <- readRDS("~/SBWD_data_analysis/EditedData/Communities/diversity/diversity_16s.RDS")

##correlation between mean richness and mean cumulative defoliation
diversity_16s_wide <- diversity_16s %>%
  pivot_wider(names_from = diversity_metric, values_from = diversity)

diversity_16s_summary <- diversity_16s_wide %>%
  group_by(site, year) %>%
  summarize(Shannon_mean=mean(Shannon), InverseSimpson_mean =mean(InverseSimpson), richness_mean=mean(richness), Evenness_mean=mean(Evenness)) 

#need mean richness per site
richness_16s <- diversity_16s %>%
  filter(diversity_metric == "richness") %>%
  dplyr::select(-4 )%>%
  dplyr::rename(richness = diversity)

mean_richness_16s <- richness_16s %>%
  group_by(site, year) %>%
  summarise(mean_richness = mean(richness))

richness_16S_cumul_HAiFLO <- mean_richness_16s %>%
        inner_join(HAIFLO, by = c("site", "year"))

#shannon and cumulative
diversity_16S_cdefo <- diversity_16s_summary %>%
        inner_join(HAIFLO, by = c("site", "year"))

#all the correlations in one plot
diversity_16S_cdefo <- diversity_16S_cdefo %>%
  dplyr::rename("Shannon Diversity" = Shannon_mean, "Richness" = richness_mean, "Evenness" = Evenness_mean)

diversity_defo_long <- diversity_16S_cdefo %>%
  pivot_longer(cols = 3:6,
                names_to = c("diversity_metric"),
                values_to = "diversity")

p5 <- diversity_defo_long  %>%
  filter(diversity_metric != "InverseSimpson_mean") %>%
ggscatter(x = "cdefo", y = "diversity", color = "year", size = 5,
          #add = "reg.line", conf.int = TRUE, 
          cor.coef = FALSE, cor.coef.size = 6, cor.method = "pearson",
          xlab = "Cumulative Defoliation", ylab = "Bacteria Diversity") + stat_cor(p.accuracy = 0.01, r.accuracy = 0.01, aes(size = 6, label = paste(..r.label.., ..p.label.., sep = "~`,`~")), show.legend = FALSE, size = 6)  + scale_color_brewer(palette = "Dark2", guide = "none") + scale_fill_brewer(palette = "Dark2", guide = "none") + 
facet_grid(diversity_metric~year, scales = "free_y") + theme_bw(base_size = 20)  + scale_x_continuous(labels = seq(0,15,5), breaks = seq(0,15,5))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))


#ITS
#need mean richness per site
richness_its <- diversity_its %>%
  filter(diversity_metric == "richness") %>%
  dplyr::select(-4) %>%
  dplyr::rename(richness = diversity)

mean_richness_its <- richness_its %>%
  group_by(site, year) %>%
  summarise(mean_richness = mean(richness))

mean_richness_its_years <- richness_its %>%
  group_by(year) %>%
  summarise(mean_richness = mean(richness))

richness_its_cumul_HAiFLO <- mean_richness_its %>%
        inner_join(HAIFLO, by = c("site", "year"))

##ITS Plot
##correlation between mean richness and mean defoliation
p6 <- richness_its_cumul_HAiFLO %>%
  ggscatter(x = "cdefo", y = "mean_richness", color = "year", size = 5,
            #add = "reg.line", conf.int = TRUE, 
            cor.coef = FALSE, cor.coef.size = 6, cor.method = "pearson",
            xlab = "Cumulative Defoliation", ylab = "Fungi Richness") + stat_cor(label.y = 450, p.accuracy = 0.01, r.accuracy = 0.01, aes(size = 6, label = paste(..r.label.., ..p.label.., sep = "~`,`~")), show.legend = FALSE, size = 6)  + scale_color_brewer(palette = "Dark2") + scale_fill_brewer(palette = "Dark2") + facet_wrap(~year) + theme_bw(base_size = 20)  + theme(legend.position = "bottom") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))

library(ggpubr)
ggarrange(p5, p6,
          heights = c(6,3),
          labels = c("A", "B"),
          font.label=list(size = 22, color = "black", face = "bold"),
          ncol = 1, nrow = 2)
```

# Fig 10 Algal Biomass

```{r}
#plot just the 3 types of algae for SI 
pp <- defo_algae_long  %>%
  dplyr::filter(algal_parameter != "Total Algae") %>%
        ggscatter(x = "cdefo", y = "algal_biomass", color = "Year", shape = "Region", 
                  size = 5, add = "reg.line", conf.int = TRUE,
                  show.legend.text = FALSE) + stat_cor(label.y.npc=1, p.accuracy = 0.01, r.accuracy = 0.01, aes(size =5, label = paste(..r.label.., ..p.label.., sep = "~`,`~")), show.legend = FALSE, size = 5) + scale_color_manual(values = c("#1b9e77", "#7570b3")) + scale_fill_manual(values = c("#1b9e77", "#7570b3")) + facet_grid(algal_parameter~Year, scales = "free_y") +xlab("Cumulative Defoliation") +ylab("Algal Biomass (ug/cm2)") + theme_bw(base_size = 16) + theme(legend.position = "bottom") + theme(legend.position = "bottom", panel.grid.major = element_blank(),
panel.grid.minor = element_blank()) + theme(legend.box.background=element_rect(color = "black", size = 1))

pp
```


# Fig S11 Extracellular Enzyme Assays
```{r}
#SI
p1 <- enzymes_corrected_summary %>%
  filter(enzyme %in% c("XY", "NAG")) %>%
        ggscatter(x = "cdefo", y = "mean_activity_corrected", color = "enzyme",
                  conf.int = FALSE, size = 4,  font.label = c(14, "plain"),
                  cor.coef = TRUE, cor.method = "pearson",cor.coef.size = 5,
                  xlab = "Cumulative Defoliation", ylab = "Mean Enzyme Activity (Weight Corrected)") +
facet_grid(enzyme~year, scales = "free_y")

p1 + theme_bw(base_size = 22) + theme(legend.position = "bottom", panel.grid.major = element_blank(),
panel.grid.minor = element_blank()) + theme(legend.box.background=element_rect(color = "black", size = 1)) + scale_fill_brewer(palette = "Dark2", guide = "none")  + scale_color_brewer(palette = "Dark2", guide = "none")
```


# Fig S12 Leaf Litter Decomposition 

```{r}
leafpacks <- readRDS("~/GitHub/SBWD_master/data/leafpacks.RDS")

#select only columns needed for decomposition analyses
lp_decomp <- leafpacks %>%
  filter(year %in% c("2019", "2020", "2021")) %>%
  dplyr::select(site, year, rep, inc_days, mass.loss.g, percent.loss)

#remove the one leafpack that was destroyed
lp_decomp <- lp_decomp[-216, ] 

lp_decomp$site <- lp_decomp$site %>%    
        factor(levels = c("U01", "U02", "U03", "C04", "C05", "C06", "C07", "L08", "L09", "L10", "L11", "L12"))

lp_decomp$year <- as.factor(lp_decomp$year)


#boxplot
p1 <- lp_decomp %>%
        ggplot(aes(x = site, y = percent.loss, fill = year)) +
        facet_wrap(~year) +
        geom_boxplot() +
        labs(title = "Leafpack Decomposition (% Mass Loss)") + theme_bw(base_size = 14) + theme(legend.position="bottom")                                             
#scatterplot against incubation time
p2<- lp_decomp %>%
        ggplot(aes(x = inc_days, y = percent.loss, color = year)) +
        geom_point(size=4)+
        labs(title = "Leafpack Decomposition (% Mass Loss) vs Incubation Time", x = "Incubation Time (days)", y = "Mass loss (%)") + theme_bw(base_size = 14)

#need to do mass loss per day
lp_decomp$inc_days <- as.numeric(lp_decomp$inc_days)
lp_decomp$percent.loss.day <- lp_decomp$percent.loss/lp_decomp$inc_days

lp_decomp %>%
        ggplot(aes(x = site, y = percent.loss.day, fill = year)) +
        facet_wrap(~year) +
        geom_boxplot() +
        labs(title = "Leafpack Decomposition (% Mass Loss per day)", y = "Percent Mass Loss per day") + theme_bw(base_size = 14) + theme(legend.position="bottom")  

#group by site to see if any differnces between years
p3 <- lp_decomp %>%
        ggplot(aes(x = year, y = percent.loss.day, fill = year)) +
        facet_grid (~site, switch = 'x') +
        geom_boxplot() +
        labs(title = "Leafpack Decomposition (% Mass Loss per day)", y = "Percent Mass Loss per day", x = "site") + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), text = element_text(size = 14))
  
#need average decomp per site per year

decomp_summary <- lp_decomp %>%
  group_by(site, year) %>%
  summarize(mean_percent_loss =mean(percent.loss), N=length(percent.loss), sd=sd(percent.loss), se = sd/sqrt(N))

decomp_summary2 <- lp_decomp %>%
  group_by(site, year) %>%
  summarize(mean_percent_loss_day =mean(percent.loss.day))

#cumulative defo

decomp_cdefo2 <- decomp_summary2 %>%
        inner_join(HAIFLO, by = c("site", "year"))

#plot correlations
p2 <- decomp_cdefo2 %>%
        ggscatter(x = "cdefo", y = "mean_percent_loss_day", color = "year",
                  conf.int = FALSE, size = 4,
                  cor.coef = FALSE, cor.method = "pearson") + 
                  xlab("Cumulative Defoliation") + ylab("% Mass Loss per Day") +
facet_wrap(~year) + stat_cor(p.accuracy = 0.01, r.accuracy = 0.01, aes(size =5, label = paste(..r.label.., ..p.label.., sep = "~`,`~")), show.legend = FALSE, size = 5) + scale_color_brewer(palette = "Dark2") + scale_fill_brewer(palette = "Dark2", guide = "none")

p2 + theme_bw(base_size = 18) + theme(legend.position = "bottom", panel.grid.major = element_blank(),
panel.grid.minor = element_blank())

```

# Fig S13 PiCRust Functional Groups

```{r}
#SI plot
       pathway_defo_all4 %>%
  filter(pathway_type %in% c("Alcohol-Degradation", "Aminoacyl-tRNAs-Charging")) %>%
       ggscatter(x = "cdefo", y = "sum_types", color = "year", 
                add = "reg.line", conf.int = TRUE, cor.coef.size = 5) +
                stat_cor(p.accuracy = 0.01, r.accuracy = 0.01,
           aes(size = 5, label = paste(..r.label.., ..p.label.., sep = "~`,`~")), show.legend = FALSE, size = 5) + xlab("Cumulative Defoliation") + ylab("Relative Abundance")  + facet_grid(pathway_type~year, scales = "free_y") + theme_bw(base_size = 18) + theme(legend.position = "bottom", panel.grid.major = element_blank(),
panel.grid.minor = element_blank())  + scale_fill_brewer(palette="Dark2") + scale_color_brewer(palette="Dark2")
```


# Fig S14 Funguild Functional Groups


```{r FunGuild, results = 'hide'}
#pull in the FunGuild output (add to repo upon acceptance)
FunGuild <- readRDS("~/SBWD_data_analysis/Code/Microbe Code/functions/FunGuild.RDS")

## How many identified 
sum(!FunGuild$guild == "na")/nrow(FunGuild) ## ~60% coverage - not bad!

#filter down to only indentifed
FunGuild <- FunGuild %>%
  filter(guild != "na")

## How many highly probable
sum(FunGuild$confidenceRanking == "Highly Probable")/nrow(FunGuild) ## 21%
#how many probable
sum(FunGuild$confidenceRanking == "Probable")/nrow(FunGuild) ##63%

#how many possible
sum(FunGuild$confidenceRanking == "Possible")/nrow(FunGuild) ##16%

#restrict to only probable and highly probable
FunGuild <- FunGuild %>%
  filter(confidenceRanking != "Possible")
#4381 ESVs out of 8464 total are identified to probable and highly probable, %52 of dataset

FunGuild_long <- FunGuild %>% 
        pivot_longer(
                cols = 28:241,
                names_to = c("site", "rep", "year"),
                names_sep = "_",
                values_to = "Raw_Abundance")

#abundance per year
FunGuild_2019 <- FunGuild_long %>%
  filter(year=="2019")

sum(FunGuild_2019$Raw_Abundance) #~644 000

#abundance per year
FunGuild_2020 <- FunGuild_long %>%
  filter(year=="2020")

sum(FunGuild_2020$Raw_Abundance) #~694 000

#abundance per year
FunGuild_2021 <- FunGuild_long %>%
  filter(year=="2021")

sum(FunGuild_2021$Raw_Abundance) #~977 000

#need to get relative abundance not raw abundance
#need to get total number of reads per site per rep
totalabund2 <- FunGuild_long %>%
  group_by(site, rep, year) %>%
  summarize(totalreads = sum(Raw_Abundance))

#join together so we can then calculate proportion
GuildAbund <- FunGuild_long %>%
        inner_join(totalabund2, by = c("site", "year", "rep"))
        
#add relative abundance column
guild_reps <- GuildAbund %>%
  mutate(Relative_Abundance = Raw_Abundance/totalreads)

guild_reps$site <- guild_reps$site %>%    
        factor(levels = c("U01","U02","U03","C04","C05", "C06","C07","L08","L09", "L10","L11","L12"))

guild_reps$year <- as.factor(guild_reps$year)

##split the complex guilds

#split the rows
guilds_sep <- guild_reps %>%
  separate_rows(guild, sep = "-", convert = TRUE) 

#sum the relative abundance 
guilds_sep_summary <- guilds_sep %>%
  group_by(site, rep, year, guild) %>%
  summarize(sum_guildreads = sum(Relative_Abundance))

unique(guilds_sep$guild)

#remove Null
guilds_sep_summary <- guilds_sep_summary %>%
  filter(guild != "NULL") #have 33 guilds now

#scale everything
guilds_sep_scale <- guilds_sep_summary %>%
        group_by(guild) %>%
      mutate(scaled.val = scale(sum_guildreads)[,1]) %>% ## need to be scaled, since we are comparing enzymes to gene counts
      group_by()## make so all are "moved up" into positive by same amount

#join defo

all_met <- guilds_sep_scale %>%
        inner_join(HAIFLO, by = c("site", "year"))

all_met2 <- all_met 
#get rid of anything with 0 value
#all_met2 <- all_met %>%
# filter(sum_guildreads > 0)

all_met2$guild <- as.factor(all_met2$guild)

#get rid of any guild with less than 3 observations 
all_met_match <- all_met2 %>%
 group_by(year, guild) %>%
summarize(numpaths = n()) %>%
  filter(numpaths >= 3) 

all_met4 <- all_met2 %>%
  inner_join(all_met_match, by = c("guild", "year"))

all_met4$guild <- as.factor(all_met4$guild)

#all three years? #nothing in all 3 years
max_met_filter_cdefo <- all_met4 %>%
  group_by(guild, year) %>%
 mutate(diffScan = summary(lm(sum_guildreads ~ cdefo))$coefficients[2,4]) %>%
      group_by(guild) %>%
      filter(max(diffScan) < 0.05)

unique(max_met_filter_cdefo$guild)

#what about in just one year
met_filter_cdefo <- all_met4 %>%
  group_by(guild, year) %>%
 mutate(diffScan = summary(lm(sum_guildreads ~ cdefo))$coefficients[2,4]) %>%
      filter(diffScan < 0.01)

unique(met_filter_cdefo$guild) #2 guilds, algal parasite and epiphyte

met_filter_cdefo_summary <- met_filter_cdefo %>%
  group_by(guild) %>%
  summarize(totalreads = sum(sum_guildreads))

#plot these
all_met4 %>%
  filter(guild %in% c("Algal Parasite", "Epiphyte")) %>%
       ggscatter(x = "cdefo", y = "sum_guildreads",
                conf.int = FALSE, cor.coef.size = 5) +
                stat_cor(p.accuracy = 0.01, r.accuracy = 0.01,
           aes(size = 5, label = paste(..r.label.., ..p.label.., sep = "~`,`~")), show.legend = FALSE, size = 5) + xlab("Cumulative Defoliation") + ylab("Relative Abundance") + facet_grid(guild~year, scales = "free_y") + theme_bw(base_size = 22) + theme(legend.position = "bottom", panel.grid.major = element_blank(),
panel.grid.minor = element_blank()) + theme(legend.box.background=element_rect(color = "black", size = 1)) + geom_smooth(data = met_filter_cdefo, method = "lm", color = "#FF0000", fill = "#FF0000") + geom_point(data = met_filter_cdefo, color = "#FF0000", fill = "#FF0000", size = 2.1)

```

# Fig S15 NO2-NO3 Concentrations

```{r nutrients}
#uses waterchem file created from streams in main rmd for this manuscript

outline <- data.frame( 
  site = c("C04", "C05", "C06", "C07", "L08", "L09", "L10", "L11", "L12", "U01", "U02", "U03"), 
  outline_color = c('black', 'black', 'black', 'black', 'black', 'black', 'black','black', 'black', 'black', 'red', 'black') 
) 

# Points defining square region for background 
square <- with(waterchem, data.frame( 
  x = c(-Inf, Inf, Inf, -Inf), 
  y = c(-Inf, -Inf, Inf, Inf)
  ))

waterchem %>%
        ggplot(aes(x=year, y=NO2.NO3.mgL, fill=year)) + 
   geom_polygon(aes(x = x,y = y, color = outline_color, fill = NA), data = merge(outline, square)) + 
        geom_boxplot() + 
        facet_grid(~site, switch = 'x') +
        xlab("Site") +
  ylab(bquote(NO[2]-NO[3] (mg/L))) + scale_color_identity() + scale_fill_identity() +
      theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), text = element_text(size = 18), panel.grid.major = element_blank(), panel.background = element_blank(),
panel.grid.minor = element_blank(), legend.position = "bottom") + scale_fill_brewer(palette="Dark2")

#TN
waterchem %>%
        ggplot(aes(x=year, y=TN.mgL, fill=year)) + 
        geom_boxplot() + 
    geom_polygon(aes(x = x,y = y, color = outline_color, fill = NA), data = merge(outline, square)) + 
        facet_grid(~site, switch = 'x') +
        xlab("Site") + scale_color_identity() + scale_fill_identity() +
      theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), text = element_text(size = 18), panel.grid.major = element_blank(), panel.background = element_blank(),
panel.grid.minor = element_blank(), legend.position = "bottom") + scale_fill_brewer(palette="Dark2")

#TP
waterchem %>%
        ggplot(aes(x=year, y=TP.mgL, fill=year)) + 
        geom_boxplot() + 
        facet_grid(~site, switch = 'x') +
        xlab("Site") +
  theme_bw() +
      theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), text = element_text(size = 18), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), legend.position = "bottom") + scale_fill_brewer(palette="Dark2")

#remove the 2 outliers in TP 
waterchem1 <- waterchem %>%
slice(-c(105, 120))

waterchem1 %>%
        ggplot(aes(x=year, y=TP.mgL, fill=year)) + 
        geom_boxplot() + 
        facet_grid(~site, switch = 'x') +
        xlab("Site") +
  theme_bw() +
      theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), text = element_text(size = 18), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), legend.position = "bottom") + scale_fill_brewer(palette="Dark2")

```
